### 2020-12-16 latex/sections/discussion.tex
latex updates
yngvi+james comments fixed

\ifdraft{Here I will discuss the myriad of fields research like this can help
for instance for health organizations, patients, taxpayers, and pharma}
hey

@@

\ifdraft{summarize the work}
hey

@@

\ifdraft{conclude the work, discuss the significance of the work}
hey

>>>>

This chapter discusses the results and the future work that could span from this research. Additionally, we conclude the work.

@@

As seen in Chapter \ref{cha:results}, our trained agent does indeed quickly rise to using the higher-level concepts, thus answering our research question \ref{rq:1}. The results are promising, as we see a popular concepts like Lorentz-Horey score rise in emphasis for the neural network, and a simple but effective concept like Material Advantage rise slowly but not exessively. We find the Unity concept to be disappointing as stated earlier in this paper; maintaining closeness of your pawns tends to be preferential. And lastly, for a the concept Aggressiveness, the fact that it falls off early is what we expected. It would have been the preferred result if we saw a greater variability in the concepts, but this could be a result from not enough training iterations, or not a complex enough model. Considering that the most emphazised HLC is the Lorentz-Horey score, and that the least is Unity, we can answer our research question \ref{rq:2}. As we know from the literature, Lorentz-Horey is considered the leading heuristic for Breakthrough, and that HLC is highly emphazised by the neural network. Secondly, the Unity heuristic is not a validated heuristic, moreso, an attempt at creating a simple heuristic that closely resembles how tight-packed your pawns are. However, the fact that the resulting graphs are relatively steady the results are not as strong as we would have hoped.

Importantly, these concepts are only a few possibilities of an infinite set of concepts, and the neural network could be learning a completely different concept then those that we tested. However, that was not the focus of this research, it was to examine if a neural network does move towards HLC's that we humans use in games.

\section{Future Work}

To iterate on this research, training a neural network with greater computing power, or with a larger neural network architecture, will allow the us to hopefully achieve a super-human neural network in Breakthrough. This would give greater confidence in the learned concepts, possibly allowing for pedagogical research on the topic regarding which concepts are optimal to train people in the game environment.

The work done in this thesis was only a proof of concept of the process of explaining the internal concept a neural network uses while acting in a board game environment. Ideally, one would like to extend this work to a larger set of board games.

Additionally applying this method to a more diverse set of environments would be an interesting field of research. We envision a self driving car agent being examined with respect to a higher level concept of aggressiveness, or a mortgage agent being examined with respect to a higher level concept of gender bias.

@@

The work done in this paper is only a first step in examining working agents in active environments with respect to human level concepts. If improved it could have a great impact in our trust on neural network that improve our daily lives. While the testbed for the research was a discrete game environment, there is a clear way forward to dynamic systems with discrete human level concepts. Our results show that for an agent that clearly doesn't achieve human level performance its actions are often selected with respect to human level concepts.

### 2020-12-16 latex/sections/introduction.tex
yngvi+james comments fixed

The world of deep learning (DL) is exciting. We have models that can examine
images and reliably identify their content. Deep learning models have
outperformed Ophthalmologists in identifying diabetic retinopathy, and identified
cancer cells where others have not. Deep learning models are used for the bleeding
edge of protein folding, to gain further understanding of the underlying structure of organisms, such as, viruses.\cite{deepmind:alphafold}
These models have done these things extremely accurately, cheap, and fast.

DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search was used to defeat the standing world champion Lee Sedol\cite{silver:alphagozero} in the incredibly expansive game of Go. This was done in the year 2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself.
These results imply that given enough time to learn the computer agent can gain a deep insight into how the game should be played.

The field of examining deep learning models to gain a richer understanding
of how they work, and what makes them so much better than humans at various tasks
is still new. This is the field of Explainable Artificial Intelligence (XAI).
If we wish to continue using artificial intelligence (AI) and machine learning (ML)
to improve our lives we must examine how they work, not only to improve
the models themselves but also to augment our own ability in many cases.
Furthermore, these explanations are a legal requirement now in many areas\cite{legalexplanation:goodman}, namely, legal, and
medical. Recently, XAI has generally been focused on Model Interpretability (MI), in
particular, interpreting the model in such a way that we might predict what the model will
do given a particular input. Another class of XAI would be Model Explainability (ME), where we can adequately explain how the internal mechanics of the model works.
A simple example of when we're able to apply ME would be when we print out a decision tree
that is generated from an ML algorithm, with the decision tree nodes and is conditions we can see exactly what the tree will do with a given input. While in MI we would have to predict the results, without going into details as to why.

In this thesis, we take a look at how we can examine an artificial neural network (ANN)
to understand which higher-level concepts (HLC) it deems important for a given state
within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
complicated like Chess or Go.

The method of examining HLC's within games has generally been done by examining the
current state of the game by evaluating them concerning some heuristic. A heuristic within games are evaluations of the end reward for a state given only the state, for example,
the number of pieces left within a game of chess. Intuitively, the amount of pieces
left is a good estimate for a state in chess, this is a higher-level concept we
use to evaluate a chess position. The piece amount can be considered as a lower-level concept
than other concepts we use. For instance, grandmaster chess players evaluate a position w.r.t. states
where the king is safe from attack or the structure of the pawn positions.
This paper examines the evaluation of a neural network of a state regarding those higher-level concepts if human players evaluate the king's safety of a state
low but the neural network highly values it, and the neural network plays better
than the player. There could be an avenue for us to improve our game by considering king safety more thoroughly.

We define the research questions:

\begin{enumerate}
      \item Given a NN that plays Breakthrough, can we evaluate if that NN recognizes the
            HLCs that human players use.
      \item Does a NN that trains itself using self-play learn some HLCs over time, and does it start to emphasize
            the HLCs that are generally considered better, more as it trains more, and oppositely does it start to
            demphasize HLCs that are considered worse.
\end{enumerate}

\subsection{Summary of the thesis}

In this thesis, we take an example game of Breakthrough, train a neural network to play the game using only self-play. That is, the neural network is trained only by playing against itself with no outside input other than the rules of the game itself. And we examine the neural networks against popular Higher-level concepts (heuristics) that we use to play the game.

>>>>

The world of deep learning (DL) is exciting. We have models that can examine images and reliably identify their content\cite{simonyan:imagerecognition}. Deep learning models have outperformed ophthalmologists in identifying diabetic retinopathy\cite{brown:retinopathy} and identified cancer cells where others have not\cite{kourou:cancer}. Deep learning models are used for the bleeding edge of protein folding, to gain further understanding of the underlying structure of organisms, such as viruses\cite{deepmind:alphafold}. These models have do these things accurately, cheap, and fast

DL in conjunction with a randomized state-space search algorithm, Monte-Carlo Tree Search, was used to defeat the standing world champion Lee Sedol in the incredibly expansive game of Go\cite{silver:alphagozero}. The researchers used reinforcement learning (RL), where the agent learned by only playing against itself. These results imply that given enough time to learn the computer agent can gain a deep insight into how a game should be played.

The field of examining deep learning models to gain a richer understanding of how they work, and what makes them so much better than humans at various tasks, is still new. This is the field of Explainable Artificial Intelligence (XAI). If we wish to continue using artificial intelligence (AI) and machine learning (ML) to improve our lives, we must examine how they work. Not only to improve the models themselves but also to augment our ability in many cases. Furthermore, these explanations are a legal requirement now in many areas\cite{legalexplanation:goodman}, namely, legal and medical. Recently, XAI has been focused on Model Interpretability (MI), in particular, interpreting the model in such a way that we can predict what the model will do given a particular input\cite{riberio:recentxai1}. Another class of XAI is Model Explainability (ME), where we attempt to adequately explain how the internal mechanics of the model works. A simple example of ME would be when we explore a decision tree generated by an ML algorithm. The decision tree nodes and conditions explain exactly what the tree will do with a given input. This differs from MI where we would have to predict the results, without going into details as to why.

In this thesis, we take a look at how we can examine an artificial neural network (ANN) to understand which higher-level concepts (HLC) it deems important for a given state within a game. Such a game can be simple like Tic-Tac-Toe or Breakthrough and complicated like Chess or Go.

The method of examining HLC's within games has generally been done by examining the current state of the game by evaluating them using some heuristic. A heuristic within games are evaluations of the expected end reward for a state, for example, by using the number and type of pieces left within a game of chess. Intuitively, the material advantage is a higher-level concept we can use to evaluate a chess position. Of course there are many other relevant concepts to consider, for example whether the king is safe from attack, and the pawn structure. This thesis examines the evaluation of a neural network of a given state with respect to those higher-level concepts. More precisely, we define the following research questions:

\begin{enumerate}
	\item \label{rq:1}Given a NN that plays Breakthrough, can we determine if that NN has learned the HLCs that human players use.
	\item \label{rq:2}Does a NN that trains itself using self-play learn some HLCs over time, and does it start to emphasize
	      the HLCs that are generally considered better the more it trains, and conversely, does it start to
	      demphasize HLCs that are considered worse.
\end{enumerate}

\subsection*{Summary}

In this thesis, we take an example game of Breakthrough, train a neural network to play the game using only self-play. That is, the neural network is trained only by playing against itself with no external input other than the rules of the game itself. And we examine the neural networks against popular higher-level concepts (heuristics) that we use to play the game.

### 2020-12-16 latex/sections/managerial.tex
yngvi+james comments fixed

\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
  have not already done so, \LaTeX{} the \path{instructions.tex} to
  learn how to setup your document and use some of the features.  You
  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
  There is also more information on working with \LaTeX{} at
  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
  This includes common problems and fixes.

  This page will disappear in anything other than draft mode.}{}

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this thesis to my family who while not understanding most of what I do
  always support me. The friends that still want to talk to me even though my
  schedule has rendered me unmeetupwithable. Lastly but certainly not least,

  Don't Panic.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
\coverchapter{Acknowledgements} 
\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}
\vspace{\baselineskip}

\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
  Note that it is important to acknowledge any funding that helped in the work}

This work was funded by \the\year~RANNIS grant ``Survey of man-eating Minke whales'' 1415550.
Additional equipment was generously donated by the Icelandic Tourism Board.

\coverchapter{Preface}

This dissertation is original work by the author, Sigurður Helgason.

\draftnote{The preface is an optional element
  explaining a little who performed what work.  See
  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
  for suggestions.

  List of publications as part of the preface is
  optional unless elements of the work have already been published.
  It should be a comprehensive list of all publications in which
  material in the thesis has appeared, preferably with references to
  sections as appropriate.  This is also a good place to state
  contribution of student and contribution of others to the work
  represented in the thesis.}

@@

\end{tabular}

>>>>

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this thesis to my family, who while not understanding most of what I do,
  always support me and the friends that still want to talk to me even though my
  schedule has rendered me unmeetupwithable. Lastly, my girlfriend Hulda, who always helps me get 
  through the tough times, and makes the good times even better.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
\coverchapter{Acknowledgements} 
\begin{quotation}
Throughout the writing of my thesis I always had support and assistance from the people around me.

I would first like to thank my supervisor, Yngvi Björnsson, his knowledge in this field is vast and he is always ready to share invaluable wisdom when it's needed. Your critical feedback and consistent positivity has pushed me and my work to a higher level.

Secondly, I would like to thank my colleagues, who provided stimulating conversations on the topic and a helpful eye from a different perspective every time I needed it.

\end{quotation}
\vspace{\baselineskip}

@@

\end{tabular}
\overfullrule=0pt

### 2020-12-16 latex/sections/methods.tex
latex updates
yngvi+james comments fixed

\chapter{Methods\label{cha:methods}}

In this chapter, we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game. The training process uses MCTS, as described in Section \ref{sec:mcts}, to guide its training. We outline the architecture used in the neural network.

@@

For a neural network to be able to use input, that input needs to be numeric. We model the board with a single $NxMx3$ array with two values $\{0,1\}$, where the first two indices on the $z$ axis represent the players, and the last layer on the $z$ axis represents which players turn it is. The first $Z$ index represents whites pawns, that is, each $x,y$ position on the board where white has a pawn that cell $x,y,0$ has the value $1$ otherwise it has $0$, and similarily on the second layer for the black player. The last $z$ layer, contains all $1$'s if it is the white players turn otherwise $0$.

One of the outputs of the nerual network is the policy vector, the policy vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$. The directions are as follows $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, and $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally. The cell values indicate the probability distribution of the neural network selecting that move. Importantly, before we select a move from the neural networks prediction we zero out the illegal moves, and renormalize the array such that it sums to $1$.

@@

The neural network architecture we opted to use was a single convolutional layer with a ReLU activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization, and a ReLu activation. Lastly, a split policy/value head, where the output of the last residual layer is split into two different outputs. The policy head are two layers: first a final convolutional layer, and then a fully connected layer with a \textit{log softmax} layer. The value head consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function. Figure~\ref{fig:nnarch}. depicts the architecture. Using this architecture we end up with $1737986$ trainable parameters.

@@

We require a global view of the board. This is why we use multiple residual layers with convolution. These residual layers stack convolutions on each other making the final convolution represent the locality of all the other localities, allowing the neural network to have a representation of the whole game state in its parameters.

The selection of parameters was done for these reasons as well as to most closely resemble the architecture described in AlphaZero.

@@

The residual layers as a concept remember the output of the previous layer and add together the results of the residual layer and the previous layer. On a higher level, this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t. two areas that are far away from each other maintain the same level of locality as two that are close.

Why we use a residual layer for applying a convolution layer multiple times is to gain locality of the whole game board. A residual layer applies a function like this $res(x) = x + l(x)$ where $l(x)$ is the function of the layer, commonly a convolutional layer.

@@

The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backward for the second player). The vector is then masked s.t. the values representing moves that can not be played on the board are given the value of $0$.

@@

\subsection{Implementation}

@@

\begin{algorithm}[t]
    \caption{Neural network compete pseudo-code}
    \label{alg:competition}
    \begin{algorithmic}[1]
        \STATE{\textbf{Input:} neural network nn1, neural network nn2}
        \STATE{\textbf{Output:} win rate for white player}
        \STATE{white\_wins $= 0$}
        \STATE{game = \textbf{initial\_breakthrough}()}
        \FOR{1 \TO 100}
        \WHILE{\textbf{not} game.is\_termina()}
        \STATE{nn1.make\_move()}
        \STATE{nn2.make\_move()}
        \ENDWHILE
        \IF{game.white\_wins()}
        \STATE{white\_wins++}
        \ENDIF
        \ENDFOR
        \STATE{\textbf{return} white\_wins/100}
    \end{algorithmic}
\end{algorithm}

@@

\section{Explainable state representations}

In this section we describe the process of trainin a Concept Activation Vector in order to linearly separate each state into points in a space with the concept and ones without the concept.

\subsection{Testing With Concept Activation Vectors}

@@

To examine the higher-level component we take a look at the internal state of the neural network itself. To do this, we take a state, and run it through the neural network, and while the state is propagating through the network we select a layer to split the network, we call that layer $l_{split}$. At that point in time the state is a mutated vector $l_{split}(l_{split-1}(...l_0(state)))$ of the initial state. We save that vector as a point in the N-dimensional space that it represents, and once we've gathered enough points in that N-dimensional space we're able to train a linear classifier using the HLC as a label. We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyper-plane that splits the space into two binary states, one that contains the HLC and another that doesn't.

@@

An example of how this would look with only two dimensions is shown in Figure \ref{fig:scattersplit}. Once we've successfully trained this linear classifier we can run another state $s$ through the neural network. Once the prediction has finished, we view the selected action $a$ of $s$ from the policy vector $p$ from the neural network. We view the loss of that selected action $a$. We then apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there. If that gradient moves in the direction of the HLC we say that the state includes the HLC, and that it doesn't if the gradient moves away from the HLC.

@@

@@

In this chapter we described the architecture of the nerual network, including how we model Breakthrough to be able to learn how to play the game. We described the processes we use to learn completely by self play. In the following chapter we will evaluate the neural network.

>>>>

\chapter{Methods}

\label{cha:methods}

In this chapter, we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game. The training process uses MCTS, as described in Section \ref{sec:mcts}, to guide its training. We outline the architecture we used for our neural network.

@@

For a neural network to be able to use input, that input needs to be numeric. We model the board with a single $N\times M\times 3$ array with two values $\{0,1\}$, where the first two indices on the $z$ axis represent the players, and the last layer on the $z$ axis represents which players turn it is. The first $z$ index represents whites pawns, that is, each $(x,y)$ position on the board where white has a pawn has the value $1$ otherwise it has $0$, and similarily on the second layer for the black player. The last $z$ layer, contains all $1$'s if it is the white players turn otherwise all $0$'s.

One of the outputs of the neural network is the policy vector. The policy vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$. The directions are as follows: $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally. The cell values indicate the probability distribution of the search selecting that move. Importantly, before we select a move from the neural networks prediction we zero out the illegal moves, and renormalize the array such that it sums to $1$.

@@

The neural network architecture we opted to use was a single convolutional layer with a ReLU\cite{hinton:relu} activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization, and a ReLu activation. Lastly, a split policy/value head, where the output of the last residual layer is split into two different outputs. The policy head are two layers: first a final convolutional layer, and then a fully connected layer with a \textit{log softmax} layer. The value head consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function. Figure~\ref{fig:nnarch} depicts the architecture. Using this architecture we end up with $1{,}737{,}986$ trainable parameters.

@@

We require a global view of the board. This is why we use multiple residual layers with convolution. These residual layers stack convolutions on each other making the final convolution represent the locality of all the other localities, allowing the neural network to have a representation of the whole game state in its parameters. The selection of parameters was done for these reasons as well as to most closely resemble the architecture described in AlphaZero\cite{Silver:alphazero}.

@@

The residual layers add the output of the previous layer and the results of the residual layer. On a higher level, this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t. two areas that are far away from each other maintain the same level of locality as two that are close.

The reason for why we use a residual layer for applying a convolution layer multiple times is to gain locality of the whole game board. A residual layer applies a function like this $res(x) = x + l(x)$ where $x$ is the input and $l(x)$ is the function of the layer, commonly a convolutional layer.

@@

The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board, $h$ is the height of the board, and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backward for the second player). The vector is then masked s.t. the values representing moves that can not be played on the board are given the value of $0$.

@@

\section{Implementation}

@@

\begin{algorithm}[t]
    \caption{Neural network compete pseudo-code}
    \label{alg:competition}
    \begin{algorithmic}[1]
        \STATE{\textbf{Input:} neural network nn1, neural network nn2}
        \STATE{\textbf{Output:} win rate for white player}
        \STATE{white\_wins $= 0$}
        \STATE{game = \textbf{initial\_breakthrough}()}
        \FOR{1 \TO 100}
        \WHILE{\textbf{not} game.is\_terminal()}
        \STATE{nn1.make\_move()}
        \STATE{nn2.make\_move()}
        \ENDWHILE
        \IF{game.white\_wins()}
        \STATE{white\_wins++}
        \ENDIF
        \ENDFOR
        \STATE{\textbf{return} white\_wins/100}
    \end{algorithmic}
\end{algorithm}

@@

\section{Explainable State Representations}

In this section we describe the process of training a Concept Activation Vector in order to linearly separate each state into points in a space with the concept and ones without the concept.

\subsection{Testing with Concept Activation Vectors}

@@

\begin{algorithm}[t]
    \caption{CAV creation pseudo-code}
    \label{alg:CAVcreation}
    \begin{algorithmic}[1]
        \STATE{\textbf{Input:} neural network nn, column\_name, breakpoint, data\_set, sample\_size}
        \STATE{\textbf{Output:} A concept activation vector for the column with \textit{column\_name}}
        \STATE{positive\_set = \textbf{select\_positive\_samples}(data\_set, column\_name, sample\_size)}
        \STATE{negative\_set = \textbf{select\_negative\_samples}(data\_set, column\_name, sample\_size)}
        \STATE{labelled\_set = positive\_set + negative\_set}
        \STATE{unlabelled\_set = dataset - unlabelled\_set}
        \STATE{model = \textbf{StochasticGradientDescentClassifier}()}
        \FOR{state \textbf{in} labelled\_set}
        \STATE{//Modify labelled dataset to contain internal representations}
        \STATE{state = nn.\textbf{get\_internal\_representation}(state)}
        \ENDFOR
        \STATE{train\_set, test\_set = \textbf{train\_test\_split}(labelled\_dataset)}
        \STATE{model.\textbf{train}(train\_set)}
        \STATE{prediction\_score, roc\_auc\_score = model.\textbf{predict}(test\_set)}
        \STATE{\textbf{return} model}
    \end{algorithmic}
\end{algorithm}

To examine the higher-level component we take a look at the internal state of the neural network itself. To do this, we take a state, and run it through the neural network, and while the state is propagating through the network we select a layer to split the network, we call that layer $l_{split}$. At that point in time the state is a mutated vector $l_{split}(l_{split-1}(...l_0(state)))$ of the initial state. We save that vector as a point in the N-dimensional space that it represents, and once we've gathered enough points in that N-dimensional space we're able to train a linear classifier using the HLC as a label. We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyper-plane that splits the space into two binary states, one that contains the HLC and another that doesn't. We show pseudo-code of this process in Algorithm \ref{alg:CAVcreation}.

@@

An example of how this would look with only two dimensions is shown in Figure \ref{fig:scattersplit}. Once we've successfully trained this linear classifier we can run another state $s$ through the neural network. Once the prediction has finished, we view the selected action $a$ of $s$ from the policy vector $p$ from the neural network. We then apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there. If that gradient moves in the direction of the HLC we say that the state includes the HLC, and that it doesn't if the gradient moves away from the HLC.

@@

\begin{algorithm}[t]
    \caption{Evaluate a neural network w.r.t concepts pseudo-code}
    \label{alg:CAVapplication}
    \begin{algorithmic}[1]
        \STATE{\textbf{Input:} \textbf{neural network} nn, \textbf{concept activation vector} cav}
        \STATE{\textbf{Output:} Portion of encountered states that contain the concept of the concept activation vector}
        \STATE{state\_count = 0}
        \STATE{has\_tcav = 0}
        \FOR{1 \TO 100}
        \STATE{current\_state = \textbf{Breakthrough}.initial\_state}
        \WHILE{! current\_state.\textbf{terminal}()}
        \STATE{current\_internal\_representation = \\ nn.\textbf{get\_internal\_representation}(current\_state)}
        \STATE{output = nn.\textbf{finish\_propagating}(current\_internal\_representation)}
        \STATE{gradient\_direction = \\ nn.\textbf{get\_gradient}(current\_internal\_representation, output)}
        \STATE{tcav\_score = cav $ \cdot $ gradient\_direction}
        \IF{\textbf{same\_direction}(tcav\_score, cav)}
            \STATE{has\_tcav++}
        \ENDIF
        \STATE{state\_count++}
        \STATE{current\_state = nn.\textbf{get\_next\_state}(current\_state)}
        \ENDWHILE
        \ENDFOR
        \STATE{\textbf{return} has\_tcav / state\_count}
    \end{algorithmic}
\end{algorithm}

In Algorithm \ref{alg:CAVapplication} we show pseudo-code depicting how we evaluate if a neural network recognizes a given concept. The resulting value from running the algorithm is, the proportion of states that the neural network encountered during playing against it self that contain the HLC.

@@

In this chapter we described the architecture of the neural network, including how we model Breakthrough to be able to learn to play the game, we described the processes we use to learn from self play. In the following chapter we will evaluate the neural network.
